---
title: "Using rbounds with simulated data"
author: "Thomas E. Love"
date: "`r Sys.Date()`"
output: 
    html_document:
        code_folding: show
        number_sections: TRUE
---

# Setup

```{r setup}
knitr::opts_chunk$set(comment = NA)
```

```{r}
library(Matching)
library(rbounds)
library(tidyverse)
```

```{r}
sim_obs <- read.csv("sim_obs.csv") %>% tbl_df

sim_obs
```

# Binary Outcome (1:1 Match)

```{r}
m.obj <- Match(Y = sim_obs$out1, 
               Tr = as.logical(sim_obs$treated), 
               X = sim_obs$ps, 
               M = 1, replace = FALSE)

summary(m.obj)
```

## Estimating $\Gamma$ with `binarysens`

```{r}
binarysens(m.obj, Gamma = 3, GammaInc = 0.25)
```

## The Matched Sample

```{r}
matches <- factor(rep(m.obj$index.treated, 2))
sim.matchedsample1 <- cbind(matches, sim_obs[c(m.obj$index.control, m.obj$index.treated),])

head(sim.matchedsample1)
```

```{r}
tmp <- sim.matchedsample1 %>% mutate(res = 10*treated + out1) %>%
    group_by(matches) %>%
    summarize(out.tr = out1[2], out.ctrl = out1[1]) 
```

```{r}
tmp %>% count(out.tr, out.ctrl)
```

So our 2x2 table would be:

2x2 Table         | Treated has `out1` | Treated no `out1`
----------------: | ---------------: | --------------:
Control has `out1`  | `r tmp %>% count(out.tr, out.ctrl) %>% mutate(res = 1:4) %>% filter(res == 4) %>% select(n)` | `r tmp %>% count(out.tr, out.ctrl) %>% mutate(res = 1:4) %>% filter(res == 2) %>% select(n)`
Control no `out1`   | `r tmp %>% count(out.tr, out.ctrl) %>% mutate(res = 1:4) %>% filter(res == 3) %>% select(n)` | `r tmp %>% count(out.tr, out.ctrl) %>% mutate(res = 1:4) %>% filter(res == 1) %>% select(n)`

# Continuous Outcome (1:1 Match)

```{r}
m.obj2 <- Match(Y = sim_obs$out2, 
               Tr = as.logical(sim_obs$treated), 
               X = sim_obs$ps, 
               M = 1, replace = FALSE)

summary(m.obj2)

psens(m.obj2, Gamma = 3, GammaInc = 0.25)
```

## Rosenbaum Bounds for Hodges-Lehmann Point Estimate

```{r}
hlsens(m.obj2, pr = 0.1, Gamma = 3, GammaInc = 0.25)
```